---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# formas

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/KTH-Library/formas/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/KTH-Library/formas/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of the R-package `formas` is to provide access to data from [FORMAS](https://formas.se/en/start-page.html), a Swedish government research council for sustainable development. This organization funds research and innovation, develops strategies, performs analyses and conducts evaluations. Areas of activity include the environment, agricultural sciences and spatial planning.

FORMAS provides open data through an API about projects funded starting from 2006. Data in the API is updated daily and may be used freely without fees or other restrictions. This R package uses this [documented API](https://formas.se/en/start-page/about-formas/what-we-do/open-data---api-containing-information-on-funded-projects/documentation-for-api-containing-information-on-funded-projects.html) to make data available for use from R.

## Installation

You can install the development version of formas like so:

``` r
library(devtools)
install_github("formas", dependencies = TRUE)
```

## Example

This is a basic example which shows you how to get data:

```{r example}
library(formas)
library(dplyr)

all_projects <- formas_projects()

# total number of projects
nrow(all_projects)

# available fields
names(all_projects)

# first few rows of latest changed projects
all_projects %>% 
  arrange(desc(`senastÄndrad`)) %>% 
  slice(1:3) %>% select(1:2, `senastÄndrad`, `beviljatBidrag`) %>%
  knitr::kable()

```

Details for a specific project or changes from a date can also be retrieved:

```{r}
# details for one specific identifier only
formas_project("2022-00327") %>% purrr::transpose()

# all changes since five days back
changes <- formas_projects_since(Sys.Date() - 5)

changes %>% select(1:2, `senastÄndrad`) %>% slice(1:5) %>% knitr::kable()

```

